---
title: "SampleSizeGLMM"
author: "Yuan Raymond SHANG"
date: "2023-08-02"
output: html_document
---

This markdown file helps to determine a reasonable sample size in a generalized linear mixed model.

Assuming we have some measurement y, which might be affected by 4 factors g1,g2,g3,and g4. We also want to see how y might be affected by the interactions between these factors.

In g1, there are 3 levels: 'y','o','m'

In g2, there are 2 levels: 'm','f'

In g3, there are 2 levels: 'a','b'

In g4, there are 3 levels: 'a3','a4','a34'

Our final goal is to estimate a reasonable SAMPLE SIZE, given alpha=0.95, beta=0.8, and some level of signals, randomness levels.

Of course, you could change Covariance Matrix part if your design matrix is different.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
library(simr)
simrOptions(progress=FALSE)
```

## Covariance Matrix
If the design is different, change the groups here to get the correct covars matrix here.

We are assuming we have 10 samples for each category, and this could be changed later.

```{r}
#Build covariance matrix, assuming 10 samples for each category
subj <- factor(1:10)
g1 <- c('y','o','m') #age
g2<-c('m','f') #Sex
g3 <- c('a','b') #mtDNA
g4 <- c("a3","a4","a34") #apoe

l_subj <- length(subj)
l_g1 <- length(g1)
l_g2 <- length(g2)
l_g3 <- length(g3)
l_g4 <- length(g4)

groups <- l_g1*l_g2*l_g3*l_g4
subj_full <- rep(subj, groups)

g1_full <- rep(rep(g1, each=l_subj), l_g2*l_g3*l_g4)
g2_full <- rep(rep(g2,each=l_subj*l_g1), l_g3*l_g4)
g3_full <- rep(rep(g3, each=l_subj*l_g1*l_g2), l_g4)
g4_full <- rep(g4,each=l_subj*l_g1*l_g2*l_g3)

covars <- data.frame(id=subj_full, g1=factor(g1_full), g2=factor(g2_full), g3=factor(g3_full),g4=factor(g4_full))

#total_groups <- 
covars
```

## Build GLMM model
Here, we need to define a GLMM model, set some basic values based on the understandings of the measurement:y.

In this demo example, assuming y has a basemean=5, the residual variance is 40% of the the basemean, thus res=2. And we are assuming the averageEffect between groups is 0.25. Note that this effect could be changed into a list, specifying different effect for different factor combinations. To make it simple here, I just assuming the effect are the same between any group and the baseline. The ultimate goal is to get the fixed array.

We also assume 0.5 randomness for each participants
```{r}
#Build GLMM model
#Base mean value
basemean=5
## residual variance, assuming the variance is around 40% of basemean
res <- 2
## Average Diff between groups, assuming 5% difference to the mean
averageEffect=0.25
#Effect size: This is not used, just for reference
Effectsize <- averageEffect/res

#
## Intercept and slopes for each factor in the GLMM model, assuming Full Model here
## Need to change based on your model if not Full Model.
fixed <- c(basemean, rep(averageEffect,(groups-1)))

## Random intercepts for participants, or participants in certain groups
randomIntercept=TRUE
### Set Random intercepts to be 10% of basemean
rand <- list(0.5)


if(randomIntercept){
  #http://www.bristol.ac.uk/cmm/learning/videos/random-intercepts.html
  model <- makeLmer(y ~ g1*g2*g3*g4 + (1|id), fixef=fixed, VarCorr=rand, sigma=res, data=covars)#+ (1|class/id)
  #for GLMM
  #makeGlmer(z ~ x + (x|g), family="poisson", fixef=b, VarCorr=V2, data=X) 
}else{
  model <- makeLmer(y ~ g1*g2*g3*g4 + (1|id), fixef=fixed, VarCorr=rand, sigma=res, data=covars)#+ (1|class/id)
}

#model <- makeLmer(y ~ g1*g2*g3*g4, fixef=fixed, VarCorr=rand, sigma=res, data=covars)#+ (1|class/id)

model
```

## Simulation to test
```{r}
#First simulate to get a rough idea
sim_treat <- powerSim(model, nsim=100, test = fcompare(y~g1))

sim_treat
```

## Samplesize estimation from simulation
```{r}
model_ext_subj <- extend(model, within="g1+g2+g3+g4", n=8)
p_curve_treat <- powerCurve(model_ext_subj, test=fcompare(y~g4), within="g1+g2+g3+g4", breaks=c(3,4,5,8))
plot(p_curve_treat)
```

